Automatically Exploring Hypothesis Spaces

### 2018-02-22 topics.tex
small amount of writing

\title{Automatically generating alternative hypotheses and controls}
%\author{Kevin Ellis, Timothy O'Donnell}

@@

@@

The fundamental hypothesis underlying AGL research is that
artificial grammar learning must engage some shared resource with first language acquisition --
and so we can gain insights into first language acquisition by
instead studying AGL in controlled experiments.
Our contribution here is the demonstration that
a single BPL model can explain how a learner can quickly infer a grammar represented as SPE--style rules.
We collected grammars from AGL datasets~\cite{gerken2010infants,marcus1999rule,frank2011three}
and presented our BPL learner with data drawn from these grammars (Table~\ref{artificialGrammarTable}),
finding that our model automatically discovers the grammars that
 underly these data sets.
\begin{table}[h!]
\centering\begin{tabular}{lcll}
  \toprule
  Grammar&Example input to learner&MAP grammar&Natural language analogues\\\midrule
  \begin{tabular}{l}
    ABA (same/different/same)\\
  Marcus et al. 1999.     
    \end{tabular}&
  \begin{tabular}{l}
    wofewo\\
    lovilo\\
    fimufi
  \end{tabular}&
  \begin{tabular}{l}
$\varnothing\to \sigma_i$\verb| / #_|$\sigma \sigma_i$
  \end{tabular}& Reduplication (eg, Tagalog)
  \\\midrule
  \begin{tabular}{l}
      ABB (different/same/same)\\
  Marcus et al. 1999. 
    \end{tabular}&
  \begin{tabular}{l}
        wofefe\\
    lovivi\\
    fimumu
  \end{tabular}&
  \begin{tabular}{l}
$\varnothing\to \sigma_i$\verb| / |$\sigma_i$\verb|_#|
  \end{tabular}&
  Reduplication (eg, Tagalog)\\\midrule
  \begin{tabular}{l}
      ABx (different/different/constant)
    \end{tabular}&
  \begin{tabular}{l}
        wofeka\\
    lovika\\
    fimuka
  \end{tabular}&
  \begin{tabular}{l}
stem$ + x$
  \end{tabular}&concatenative morphology
  \\\midrule
  \begin{tabular}{l}
    AAx (same/same/constant)\\Gerken 2006. %~\cite{gerken2010infants}
    \end{tabular}&
  \begin{tabular}{l}
        wowoka\\
    loloka\\
    fifika
  \end{tabular}&
  \begin{tabular}{l}
    stem + $x$\\
$\varnothing\to \sigma_i$\verb| / #_|$\sigma_i$
  \end{tabular}&
  \begin{tabular}{l}
    reduplication\\concatenative morphology
    \end{tabular}\\\midrule
  \begin{tabular}{l}
    AxA (same/constant/same)\\Gerken 2006. %~\cite{gerken2010infants}
    \end{tabular}&
  \begin{tabular}{l}
        wokawo\\
    lokalo\\
    fikafi
  \end{tabular}&
  \begin{tabular}{l}
$x + $ stem\\
    $\varnothing$ $\to\sigma_i$  / \verb|# _| $\sigma$ $\sigma_i$
  \end{tabular}&
  \begin{tabular}{l}
    Infixing\\Reduplication
    \end{tabular}
  \\\midrule
  \begin{tabular}{l}
    Pig Latin
    \end{tabular}&
  \begin{tabular}{l}
    \textipa{pIg}$\to$\textipa{Igpe}\\
    \textipa{latIn}$\to$\textipa{atIle}\\
    \textipa{\ae sk}$\to$\textipa{\ae ske}\\
  \end{tabular}&
  \begin{tabular}{l}
      $\varnothing\to$\verb|C|$_i$ $/$ \verb|#C|$_i$\verb| [ ]* _ #|\\%
      $\varnothing\to$ \textipa{e} $/$ \verb| _ #|\\%
 \verb|C|$\to\varnothing$ $/$ \verb|# _|
  \end{tabular}  & \begin{tabular}{l}
    \textbf{Tim, is there an  analog?}\\
    \textbf{word initial segments moves to  }\\
    \textbf{word final?}
    \end{tabular}
  \bottomrule  \end{tabular}
\caption{Using BPL to model artificial grammar learning. Learner given 5 examples.}\label{artificialGrammarTable}
\end{table}

>>>>

\title{Automatically Exploring Hypothesis Spaces}
%\author{Kevin Ellis, Timothy O'Donnell}

@@

In this work we will use this probabilistic framing
to explore spaces of grammars
that an AGL learner could use to explain
the pronunciations of words.
Here the grammars will be SPE-style rewrite rules~\cite{chomsky1968sound},
commonly used in generative phonology,
but at a high level this framing can apply to
learning other aspects of grammar like semantics~\cite{piantadosi2011learning}, syntax~\cite{perfors2011learnability}, and morphology~\cite{tim}.

@@

The fundamental hypothesis underlying AGL research is that
artificial grammar learning must engage some shared resource with first language acquisition --
and so we can gain insights into first language acquisition by
instead studying AGL in controlled experiments.
Our contribution here is the demonstration that
a single BPL model can explain how a learner can quickly infer a grammar represented as SPE--style rules.
We collected grammars from AGL datasets~\cite{gerken2010infants,marcus1999rule,frank2011three}
and presented our BPL learner with data drawn from these grammars (Table~\ref{artificialGrammarTable}),
finding that our model automatically discovers the grammars that
 underly these data sets.
\begin{table}[h!]
\centering\begin{tabular}{lcll}
  \toprule
  Grammar&Example input to learner&Grammar we infer&Natural language analogues\\\midrule
  \begin{tabular}{l}
    ABA (same/different/same)\\
  Marcus et al. 1999.     
    \end{tabular}&
  \begin{tabular}{l}
    wofewo\\
    lovilo\\
    fimufi
  \end{tabular}&
  \begin{tabular}{l}
$\varnothing\to \sigma_i$\verb| / #_|$\sigma \sigma_i$
  \end{tabular}& Reduplication (eg, Tagalog)
  \\\midrule
  \begin{tabular}{l}
      ABB (different/same/same)\\
  Marcus et al. 1999. 
    \end{tabular}&
  \begin{tabular}{l}
        wofefe\\
    lovivi\\
    fimumu
  \end{tabular}&
  \begin{tabular}{l}
$\varnothing\to \sigma_i$\verb| / |$\sigma_i$\verb|_#|
  \end{tabular}&
  Reduplication (eg, Tagalog)\\\midrule
  \begin{tabular}{l}
      ABx (different/different/constant)
    \end{tabular}&
  \begin{tabular}{l}
        wofeka\\
    lovika\\
    fimuka
  \end{tabular}&
  \begin{tabular}{l}
stem$ + x$
  \end{tabular}&concatenative morphology
  \\\midrule
  \begin{tabular}{l}
    AAx (same/same/constant)\\Gerken 2006. %~\cite{gerken2010infants}
    \end{tabular}&
  \begin{tabular}{l}
        wowoka\\
    loloka\\
    fifika
  \end{tabular}&
  \begin{tabular}{l}
    stem + $x$\\
$\varnothing\to \sigma_i$\verb| / #_|$\sigma_i$
  \end{tabular}&
  \begin{tabular}{l}
    reduplication\\concatenative morphology
    \end{tabular}\\\midrule
  \begin{tabular}{l}
    AxA (same/constant/same)\\Gerken 2006. %~\cite{gerken2010infants}
    \end{tabular}&
  \begin{tabular}{l}
        wokawo\\
    lokalo\\
    fikafi
  \end{tabular}&
  \begin{tabular}{l}
$x + $ stem\\
    $\varnothing$ $\to\sigma_i$  / \verb|# _| $\sigma$ $\sigma_i$
  \end{tabular}&
  \begin{tabular}{l}
    Infixing\\Reduplication
    \end{tabular}
  \\\midrule
  \begin{tabular}{l}
    Pig Latin
    \end{tabular}&
  \begin{tabular}{l}
    \textipa{pIg}$\to$\textipa{Igpe}\\
    \textipa{latIn}$\to$\textipa{atIle}\\
    \textipa{\ae sk}$\to$\textipa{\ae ske}\\
  \end{tabular}&
  \begin{tabular}{l}
      $\varnothing\to$\verb|C|$_i$ $/$ \verb|#C|$_i$\verb| [ ]* _ #|\\%
      $\varnothing\to$ \textipa{e} $/$ \verb| _ #|\\%
 \verb|C|$\to\varnothing$ $/$ \verb|# _|
  \end{tabular}  & \begin{tabular}{l}
    \textbf{Tim, is there an  analog?}\\
    \textbf{word initial segments moves to  }\\
    \textbf{word final?}
    \end{tabular}
  \bottomrule  \end{tabular}
\caption{Using BPL to model artificial grammar learning. Learner given 5 examples.}\label{artificialGrammarTable}
\end{table}
